### 2018-10-03 tex/cover-letter/cover_letter.tex
Added the journal cover letter

\documentclass{letter}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\signature{Jonathan Ström}
\address{School of Engineering \\ Brown University \\ Providence 02912}
\begin{document}

\begin{letter}{Editor of Environmental Science \& Technology \\
1155 Sixteenth Street N.W. \\
Washington, DC 20036}
\opening{Dear Sir or Madam:}

I am pleased to submit this manuscript entitled "Modeling Variation in Indoor Air Contaminant Concentration \& Impact of Preferential Contaminant Pathways on Vapor Intrusion" for review and possible publication in Environmental Science \& Technology.
This is original work, and has not been previously submitted to any other journal or conference proceedings.

Using numerical models and published field data from two vapor intrusion (VI) sites, we explore the causes of the temporal variation in indoor air contaminant concentrations in VI impacted houses, a poorly understood phenomena, and especially the role of preferential contaminant pathways in this.

We found under which conditions one may expect to observe significant changes in indoor air concentrations, and when one should be especially be concerned with preferential pathways.
These finding help understand and explain the temporal variation that has been observed at many VI sites, and giving practical guidance for assessment of these sites, something that we expect will be of great interest to your readership.

All of the work presented is either created by us or properly referenced, including the TOC art which adhere to the TOC art policy and is the prescribed size of 240 by 135 points.

Thank you for considering this manuscript for publication.
\closing{Sincerely,}

\end{letter}
\end{document}