### 2019-05-02 tex/highlights/highlights.tex

\documentclass{article}

\usepackage[utf8]{inputenc}

\begin{document}
\section*{Highlights}

\begin{itemize}
  \item Preferential pathways may lead to the false conclusion that there are indoor sources present at a VI site.
  \item Preferential pathways can dramatically increase the role of advective transport, provided there is a permeable subslab region to facilitate this.
  \item At VI sites where advective transport potential is high, much of the observed variation may be attributed to fluctuations in building pressurization; where this potential is much lower, fluctations in air exchange rate may significantly contribute to indoor contaminant variability.
  \item In the absence of a preferential pathway, indoor contaminant concentration vary roughly by up to a factor of three across a few days; up to an order magnitude variation may be expected over a week.
\end{itemize}
\end{document}