### 2019-06-04 tex/reviewer_response/reponse.tex
Revision

\documentclass{letter}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[textwidth=6in,textheight=14in]{geometry}
\signature{Jonathan Ström}
\address{School of Engineering \\ Brown University \\ Providence 02912}
\begin{document}

\begin{letter}{ }
\opening{Dear reviewers:}

We thank the reviewers for their efforts in reviewing our paper, and have
responded to their questions and comments as outlined below.

Reviewer \#1: This paper studied temporal variations of indoor contaminant concentrations
due to Vapor Intrusion by analyzing the field data and performing simulations. Some
interesting conclusions were obtained from the analysis of data. The method seems
appropriate. Therefore, I suggest the paper to be accepted by addressing the following
concerns:

\textit{We thank the reviewer for the positive assessment of our paper.}

Comment 1. The graphic abstract should be more self-contained. Necessary explanations
are suggested.

\textit{Since the graphical abstract is not an element required by this journal, we decided that since
it seemed to raise more questions than it answered, it would be better to remove it entirely,
rather than add even more complexity to it.}

Comment 2. Please provide practical implications of the obtained conclusions. For example, how to
detect or reduce peak concentration to protect human health?

\textit{We have done so as part of a significantly expanded "Conclusions" section, and thank the
reviewer for suggesting this useful addition.}

Reviewer \#2: In my opinion it is a remarkably prepared material. During almost of 10 years of
my experience as a reviewer I have never reviewed such an outstanding manuscript, both in
terms of the structure and balanced proportion between the complexity and the way of
presentation of material as well as the clarity of narration. It should be particularly
appreciated due to the fact that presented material relates to quite complex issue like a long-
term data elaboration and interpretation, which is even more difficult taking into account all
possible factors influencing obtained results.
I could not find any lingual, grammatical or punctuation slip. I strongly recommend to accept
this manuscript for publication.

\textit{What can we say but thank you to this reviewer!}

\closing{Sincerely,}

\end{letter}
\end{document}