### 2019-06-14 tex/highlights/highlights.tex
Added new highlights

\documentclass{article}

\usepackage[utf8]{inputenc}

\begin{document}
\section*{Highlights}

\begin{itemize}
  \item Preferential pathways may lead to the false conclusion that there are indoor sources present at a VI site.
  \item Preferential pathways can dramatically increase the role of advective transport, provided there is a permeable subslab region to facilitate this.
  \item At VI sites where advective transport potential is high, much of the observed variation may be attributed to fluctuations in building pressurization; where this potential is much lower, fluctations in air exchange rate may significantly contribute to indoor contaminant variability.
  \item In the absence of a preferential pathway, indoor contaminant concentration vary roughly by up to a factor of three across a few days; up to an order magnitude variation may be expected over a week.
\end{itemize}
\end{document}

Preferential pathways may lead to the false conclusion that there are indoor sources present at a VI site.
Indoor concentrations vary roughly by a factor of three over a few days; up to an order of magnitude over a week.
Preferential pathways can dramatically increase the role of advective transport, provided there is a permeable subslab zone.
At sites with high advective transport potential, fluctuations in pressurization cause much of the observed variability.
Air exchange rate fluctuations significantly contribute with variability at sites with low advective transport potential.